Analysis of Water Security in Jordan

### 2022-04-02 outputs/paper/paper.Rmd
Datasheet completed, abstract completed, bibiliography completed.

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references_datasheet.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
**DataSheet**

Extract of the questions from @gebru2021datasheets

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the source and availability of water in different populations of Jordan.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Jonathan Goodwin
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

@@

	- Each row of the dataset gives a percent of the population 

@@

	- 

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

	- TBD

@@

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

@@

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

@@

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

@@

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

@@

>>>>

---
title: "Analysis of Water Security in Jordan"
author: 
  - Jonathan Goodwin
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "Water source and availability are important factors in determining the health and safety of a society. We obtain data regarding the source of water in Jordan and the time it takes residents to access it. We find that the most common source of water is piped into residents, and that a significant portion of Rural residents are receiving water through a tanker truck. These findings have implication for the further development of infrastrcuture in Jordan."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references.bib
---
```{r setup, include=FALSE,warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

```{r, echo = F, warning = F, message = F, results = F}
### Reading CSV ###
Data <- 
  read.csv(here::here("outputs/data/cleaned_data.csv"))
```

# Introduction

You can and should cross-reference sections and sub-sections. For instance, Section \@ref(data). R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).

# Data

  The dataset is from the Jordan Population and Family Health Survey(JPFHS). The survey was conducted by the Department of Statistics, [@citeDOS] from June 7, to October 31, of the year 1997,through the Demographic and Health Survey(DHS) program.The survey was funded by the U.S Agency for International Development, while Macro International Inc. provided technical support of the survey.

  The survey covers 7600 households, the sample is nationally represenative, the intent of the survey was to provide information on levels of fertility, infant and child mortality and family planning. Of this survey, this analysis is focused on the information regarding the source of water and the time availability of water. This information was available in the JPFHS report on page 32, table 2.7.

  To compile the data set, the R language was used, [@citeR], along with the packages, Pointblank [citePointblank], haven [citehaven], the paper [@citeknitr]

```{r, echo = FALSE}

```

Talk more about it.

Also bills and their average. (Notice how you can change the height and width so they don't take the whole page?)

```{r, echo = FALSE}

```

Talk way more about it.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Weaknesses and next steps

More information regarding the availability deficiencies would be the next steps. Are the rural areas receiving water via tanker truck completely remote or can they receive piped water sources?

\appendix

# Appendix {-}

## DataSheet

Datasheet questions from @gebru2021datasheets

@@

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to analyze the source and availability of water in different populations of Jordan.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - Jonathan Goodwin
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - No one.
4. *Any other comments?*
    - No.

@@

	- Each row of the dataset corresponds to a characteristic, with values of the percent of rural and urban residents of Jordan that fit that characteristic, with a final column giving a total for the entire population.

@@

	- There are 10 rows of data, 9 correspond to sources of water, one corresponds to the time availability of the water source.

@@

	- The dataset is all instances of those that responded to the questionaire, so it does not encompass the entire population of Jordan, the larger set. According to the DHS the sample is representative of the national population.

@@

	- Each instance consists of a numeric value for the percentage of residents which responded affirmatively to that characteristic in the survey.

@@

	- No.

@@

	- The dataset does not rely on any external sources.

@@

	- No, the data is all publicly available from the Department of statistics Jordan and required no special permissions to access.

@@

	- The sub populations are urban and rural residents of Jordan.

@@

	- No individuals can be identified from the dataset.

@@

	- None of the data is of a sensitive nature.

@@

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- The data was originally acquired by the Department of Statistics[Jordan], [@citeDOS] via the Jordan Population and Family Health Survey. A table from that report found on page 32, was then used for this analysis.
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- Subjects for the survey were interviewed in person.
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- The sample was from the Jordan population and the sampling strategy of multistage stratified design was used in collecting the sample.
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- The Department of Statistics in Jordan, as well as Marcus International Inc. under contract funded by the United States Agency for International Development.
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- The data was collected from June 7 of 1997 to October 31 of 1997.
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- DHS surveys are approved and review by the ICF Institutional Review Board( IRB).[@citeDHS_Ethics].
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- Individuals of the survey were contacted directly for the survey.
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- Participants are notified of all aspects of the 
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- According to the DHS, informed consent statements are read to respondents who may accept or decline participation.[@citeDHS_Ethics]
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- Participants of the survey are told they may terminate participation at any time.
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- The original survey was subject to ethical review this particular subset of that study has not been.
12. *Any other comments?*
	- No

@@

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- In the process of building this dataset
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- Yes, both the raw data acquired through the table, and the cleaned version of the dataset are available in the repository associated with this analysis.
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- The R language and packages associated with the cleaning process are all freely available.
4. *Any other comments?*
	- No.

@@

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- Prior to this analysis the dataset was only used as part of the original Survey.
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- Yes, it is available on github [here](https://github.com/Jon-Goodwin/Paper-4)
3. *What (other) tasks could the dataset be used for?*
	- The dataset has been minimized for this analysis but the full raw data includes other factors regarding water security that may be of interest.
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- No.
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- No.
6. *Any other comments?*
	- No.

@@

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- No.
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- The dataset is available via Github.
3. *When will the dataset be distributed?*
	- The dataset is currently available via Github.
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- No.
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- No.
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- No.
7. *Any other comments?*
	- No.

@@

1. *Who will be supporting/hosting/maintaining the dataset?*
	- The dataset will be available on Github
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- 
3. *Is there an erratum? If so, please provide a link or other access point.*
	- No.
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- No.
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- No.
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- No.
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- No.
8. *Any other comments?*
	- No.

@@

## Code
Repository associated with this analysis is available at [github](https://github.com/Jon-Goodwin/Paper-4)]
